---
title: EVALUATING THE BENEFITS, BARRIERS AND FACILITATORS OF GROUP INTERVENTIONS AMONG
  PERSONS BATTLING ALCOHOL AND DRUG ABUSE IN TURBO SUB-COUNTY, UASIN GISHU COUNTY
author: "Allan Kimaina and Julius Barasa"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  pdf_document: default
subtitle: Population Health
always_allow_html: yes
---



```{r setup, include=FALSE}
options(java.parameters = "-Xmx15g")

knitr::opts_chunk$set(warning=FALSE,
                      message=FALSE,
                      echo=FALSE,
                      #dpi=96,
                     # fig.width=7,# fig.height=4, # Default figure widths
                     # dev="png", #dev.args=list(type="cairo"), # The png device
                      # Change to dev="postscript" if you want the EPS-files
                      # for submitting. Also remove the dev.args() as the postscript
                      # doesn't accept the type="cairo" argument.
                      error=FALSE)
 
# Evaluate the figure caption after the plot
#knitr::opts_knit$set(eval.after='fig.cap')
 
# Use the table counter that the htmlTable() provides
options(table_counter = TRUE)
 
# Use the figCapNo() with roman letters
#options(fig_caption_no_roman = TRUE)
#options(kableExtra.latex.load_packages = F)

# Then install the Grmd-package by running below code:
#devtools::install_github("gforge/Grmd")

#devtools::install_github("kassambara/easyGgplot2")
library(easyGgplot2)

# function to install missing packages
ipak <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg))
    install.packages(new.pkg, dependencies = TRUE, repos='http://cran.rstudio.com/')
  sapply(pkg, require, character.only = TRUE)
}

#install.packages('package_name', dependencies=TRUE, repos='http://cran.rstudio.com/')

packages =c( "dplyr",  "readxl","Hmisc","Gmisc", "magrittr", "flextable", "MASS", "tidyverse", "caret", "knitr", "kableExtra","xtable", "stargazer", "ggpubr", "haven", "tidycomm", "gtable", "ggpmisc",  "gtsummary", "emmeans", "ggstatsplot", "jtools", "lme4", "texreg", "car", "afex", "lmerTest")

ipak(packages)


select = dplyr::select; summarize = dplyr::summarize; rename = dplyr::rename; mutate = dplyr::mutate;

source("unbalanced_functions.R")

```


# Study Background:

Use of alcohol and substances of abuse is a common health problem affecting about  10-16% of Kenyans. Most of the care is provided in inpatient rehabilitation centers which are not  accessible to a majority of the population due to the financial implications. Small community-based support groups have potential to increase accessibility of alcohol and drug abuse rehabilitation  services. As part of a care program to address the treatment gap for mental and substance, we  established 28 informal alcohol support groups in Turbo sub-county and these have been running  over the past two years. The goal of this study therefore is to objectively evaluate the benefits of these groups.  


Analysis Objective: 

1) To establish the benefits of community-based alcohol support groups on the perceived quality of  life of patients with alcohol and drug abuse disorders. 
2) To assess the benefits of community-based alcohol support groups on alcohol drinking patterns  of the patients in terms of number of drinking days and the number of drinks per drinking day 
 3) To explore the existing social support systems for alcohol and substance abuse patients that would  facilitate recovery among group members.  
4) To establish stigma and other barriers to effective participation in community alcohol support  groups.
Population & Settings

118 participants will be randomly sampled from the existing group members.
Eligibility Criteria.  

1. Persons who were screened and linked to community support groups in the sub-county.
2. A support group member who has attended at least six group meetings. 
3. Aged 18 years and above. 

Exclusion criteria 

1. Persons with alcohol use disorders within Turbo subcounty who hasn’t been enrolled into any  of the existing groups. 
2. Persons who will present during the data collection session while drunk. 
3. All members of the support groups aged below 18 years. 



```{r warning=FALSE}

describeMissing= function (x, html = TRUE, number_first = TRUE, percentage_sign = TRUE, 
    language = "en", useNA.digits = 1, ...) {
    
    if (!any(is.na(x))) 
        return(invisible())
    df_arg_list <- list(x = is.na(x), html = html, number_first = number_first, 
        percentage_sign = percentage_sign, language = language, 
        digits = useNA.digits)
    dot_args <- list(...)
    for (n in names(dot_args)) {
        if (!n %in% names(df_arg_list)) {
            df_arg_list[[n]] <- dot_args[[n]]
        }
    }
    missing <- fastDoCall(describeFactors, df_arg_list)
   rownames(missing)<-c("FALSE","Missing")
    return(missing["Missing", ])
}

desc_both <- function(x, ...) {
    result <- c(
      describeMean(x, useNA="no"),
      describeMedian(x, useNA="no"),
      describeMissing(x)
    )
    return(result)
}

desc_mean <- function(x, ...) {
    result <- c(
      describeMean(x, useNA="no"),
      describeMissing(x)
    )
    return(result)
 }
 

desc_median <- function(x, ...) {
    result <- c(
      describeMedian(x, useNA="no"),
      describeMissing(x)
    )
    return(result)
 }

 
 MainX<<-NULL
# Creating a wrapper for getting descriptive statistics
getTable1Stats <- function(x, y, cont_fx=desc_both, data=dataset, digits = 1,statistics = T,na.rm=na.rm.var, header_count = T){
   MainX <<- append(MainX, c(x))
  data=data%>%drop_na(all_of(y))
  if(na.rm){
    data=data%>%drop_na(all_of(x))
  } 
  getDescriptionStatsBy(x = data[[x]], 
                        by = data[[y]],
                        digits = digits,
                        statistics = statistics,
                        continuous_fn = cont_fx,
                        hrzl_prop = T,
                        total_col_show_perc = T,
                        header_count =  header_count )
  
}

relevelBy <- function(varList, data){
  for (factorName in names(varList)) {
    for(var in varList[[factorName]]){
     # print(var)
      data[[var]] =relevel(factor(data[[var]]),factorName)
    }
  }
  return(data)
}


getMultipleTable1Stats <- function(x, y,y2,y3, cont_fx=desc_both, data=dataset, digits = 1,statistics = T,hrzl_prop = T,na.rm=na.rm.var){
 # data=data%>%drop_na(all_of(y))
  if(na.rm){
    data=data%>%drop_na(all_of(x))
  } 
 yy1= getDescriptionStatsBy(x = data[[x]], 
                        by = data[[y]],
                        digits = digits,
                        statistics = statistics,
                        continuous_fn = cont_fx,
                        hrzl_prop = hrzl_prop,
                        total_col_show_perc = T,
                        header_count = TRUE)
 
 yy2= getDescriptionStatsBy(x = data[[x]], 
                        by = data[[y2]],
                        digits = digits,
                        statistics = statistics,
                        continuous_fn = cont_fx,
                        hrzl_prop = hrzl_prop,
                        total_col_show_perc = T,
                        header_count = TRUE)
 
 yy3= getDescriptionStatsBy(x = data[[x]], 
                        by = data[[y3]],
                        digits = digits,
                        statistics = statistics,
                        continuous_fn = cont_fx,
                        hrzl_prop = hrzl_prop,
                        total_col_show_perc = T,
                        header_count = TRUE)
  
  
  result <- tryCatch({
    return(cbind(yy1,yy2,yy3))
    
    }, warning = function(war) {
    
      # warning handler picks up where error was generated
      print(paste(x,"ERROR"))
      print(paste("MY_WARNING:  ",war))
     
    
    }, error = function(err) {
    
      # error handler picks up where error was generated
      print(paste(x,"ERROR"))
      print(paste("MY_ERROR:  ",err))
      #return(f)
    
    }, finally = {
    
     # print(paste(x,"Done"))

    }) # END tryCatch
  
}

stargazer2 <- function(model, odd.ratio = F, ...) {
  if(!("list" %in% class(model))) model <- list(model)
    
  if (odd.ratio) {
    #exponentiate <- function(x) exp(x)
    coefOR2 <- lapply(model, function(x) exp(coef(x)))
    CIOR2 <- lapply(model, function(x) exp(confint(x)))
    seOR2 <- lapply(model, function(x) exp(coef(x)) * summary(x)$coef[, 2])
    p2 <- lapply(model, function(x) summary(x)$coefficients[, 4])
    stargazer(model, coef = coefOR2, se = seOR2, p = p2,  ci.custom =CIOR2, ...)
    
  } else {
    stargazer(model, ...)
  }
}

generateInterpretation= function(outcome,topn,model.df){
  
  model.df%>%mutate(
  interpretation=ifelse( term!="(Intercept)",paste("Fixing all else constant, a unit increase in ",tolower(variable_name),", changes the odds of ",outcome," by: ", estimate_or, " (",conf.low_or, "-",conf.high_or,") on average. In other words, the odds of ",outcome," changes by", 100 * (estimate_or - 1), "% due to each unit increase in ",tolower(variable_name)),paste("Not meaningful but can be roughly translated to the odds of ",outcome," which is ",estimate_or, " (",conf.low_or, "-",conf.high_or,") ,fixing all else constant")),
  
  interpretation=ifelse(categorical==T,paste("The odds of ",outcome," in ", tolower(variable_name), " is ", estimate_or, " (",conf.low_or, "-",conf.high_or,") times higher than the odds of ", outcome," in the reference group, fixing all else constant." ),interpretation )
)%>%select(variable_name,interpretation )%>%
  kable( "html", booktabs = T, longtable = F,  digits=2,
         caption = paste("Model Interpretation | ",capitalize(outcome)) ) %>%
  kable_styling(bootstrap_options = c("striped","hold_position","condensed","responsive"))
  

}

generateLogitPlot= function(outcome,topn,model.df){
  ggplot(model.df%>%top_n(topn,p.value), aes(estimate, variable_name, color = variable_name )) +
      geom_point() + geom_errorbarh(aes(xmin = conf.low, xmax = conf.high)) + geom_vline(xintercept = 1, color = "blue", lty = 2) +
      ggtitle( sprintf( paste("Precision Comparison | ",capitalize(outcome)) ) )+  xlab("Log Odds Ratio") + ylab("Covariate")+
      theme_economist()+ guides(fill=FALSE, color=FALSE)+theme(plot.title = element_text(size=14, face="bold"))
}

```



```{r warning=F}
dataset = foreign::read.spss("TURB0 GROUP DATA ENTRY 8th Oct.sav", to.data.frame=TRUE)%>%
  mutate(
    age_cat=ifelse(age<18,"<18",NA),
                     age_cat=ifelse( age>=18&age<=25,"18 – 25",age_cat),
                     age_cat=ifelse( age>=26&age<=35,"26 – 35",age_cat),
                     age_cat=ifelse( age>35,">35",age_cat),
     ageoffirstdrink_cat=ifelse(ageoffirstdrink<18,"<18",NA),
                     ageoffirstdrink_cat=ifelse( ageoffirstdrink>=18&ageoffirstdrink<=25,"18 – 25",ageoffirstdrink_cat),
                     ageoffirstdrink_cat=ifelse( ageoffirstdrink>=26&ageoffirstdrink<=35,"26 – 35",ageoffirstdrink_cat),
                     ageoffirstdrink_cat=ifelse( ageoffirstdrink>35,">35",ageoffirstdrink_cat),
    alcoholicdrinkstakes_cat = case_when(
                                   
                                    str_detect(alcoholicdrinkstakes,"^busaa") ~ "busaa +",
                                    str_detect(alcoholicdrinkstakes,"^chang") ~ "chang’aa +",
                                    str_detect(alcoholicdrinkstakes,"^beer") ~ "beer +",
                                    str_detect(alcoholicdrinkstakes,"none") ~ "none",
                                    is.na(alcoholicdrinkstakes) ~ NA_character_,
                                    TRUE ~ "others"),
    othersubstances_cat = case_when(
                                   
                                    str_detect(othersubstances,"^cig") ~ "cigaratte +",
                                    str_detect(othersubstances,"^sig") ~ "cigaratte +",
                                    str_detect(othersubstances,"^bhang") ~ "bhang",
                                    str_detect(othersubstances,"^t") ~ "tobacco +",
                                    str_detect(othersubstances,"none") ~ "none",
                                    is.na(othersubstances) ~ NA_character_,
                                    TRUE ~ "others"),
  )%>%mutate(
                                   male=ifelse(gender=="male", 1,0),
                                   married=ifelse(maritalstatus=="married", 1,0),
                                   christian=ifelse(religion=="christian", 1,0),
                                   educated=ifelse(education=="none", 0,1),
                                   unemployed=ifelse(work=="unemployed", 1,0),
                                   domain1_post=VAR00002, 
                                   domain1_pre=VAR00004, 
                                   domain2_post=VAR00006,
                                   domain2_pre=VAR00008, 
                                   domain3_post=VAR00010, 
                                   domain3_pre=VAR00012, 
                                   domain4_post=VAR00014, 
                                   domain4_pre=VAR00016
                                )

```


# Univariate Analysis

Descriptive Analysis under the univariate setting will be applied to summarize central tendencies and distribution for metric variables. 

* Categorical variables such as gender, marital status will be summarized using frequencies and percentages. 
* Categorical variables derived from continuous variables such as  x1, x2 will be categorized using clinically acceptable limits derived from literature. 
* Continuous variables that assume the Gaussian distribution will be summarized using mean and corresponding SD.
* Continuous variables that are skewed will be summarized as median and the corresponding interquartile range (IQR). 
* The test for normality assumptions will be done using the Shapiro-Wilks,Shapiro-Francia.
* Missing values will be presented using frequency and % percentages


## Table 1: Demographics and socio-economic characteristics
```{r warning=FALSE}
na.rm.var=F;stats=T;outcome="gender";

# Getting descriptive statistics 
mergeDesc("Age (Years)" = getTable1Stats("age",outcome,desc_both,statistics=stats,na.rm=na.rm.var),
          "Age (%)" = getTable1Stats("age_cat",outcome,desc_both,statistics=stats,na.rm=na.rm.var),
         # "Gender" = getTable1Stats("gender",outcome,desc_both,statistics=stats,na.rm=na.rm.var),
          "Marital status" =getTable1Stats("maritalstatus",outcome,desc_both,statistics=stats,na.rm=na.rm.var),  
          "Religion" =getTable1Stats("religion",outcome,desc_both,statistics=stats,na.rm=na.rm.var),
          
          "Highest level of education" = getTable1Stats("education",outcome,desc_both,statistics=stats,na.rm=na.rm.var),
          "Employment Status" = getTable1Stats("work",outcome,desc_both,statistics=stats,na.rm=na.rm.var),
           "Estimated monthly income" = getTable1Stats("income",outcome,desc_both,statistics=stats,na.rm=na.rm.var),
           "Currently living with" = getTable1Stats("currentfamily",outcome,desc_both,statistics=stats,na.rm=na.rm.var),
           "Primary support system" = getTable1Stats("supportsystem",outcome,desc_both,statistics=stats,na.rm=na.rm.var)
          )%>%
   htmlTable( caption  = "<b> Demographics and socio-economic characteristics </b>",useViewer=T,ctable = TRUE, align = 'lcccc',
              n.cgroup = c(1,2,1 ), cgroup = c('', 'Gender', '') ,tfoot="<sup>&Dagger;</sup>"
              )%>% htmlTable::addHtmlTableStyle(col.rgroup = c("#FFF", "#EEF")) 



```



## Table 2: Alcohol and drug abuse characteristics
```{r warning=FALSE}
na.rm.var=F;stats=T;outcome="gender";

# Getting descriptive statistics 
mergeDesc(
           "History of alcohol in the family" = getTable1Stats("familyalcoholhistory",outcome,desc_both,statistics=stats,na.rm=na.rm.var),
           "Age of first alcoholic drink (Years)" = getTable1Stats("ageoffirstdrink",outcome,desc_both,statistics=stats,na.rm=na.rm.var),
           "Age of first alcoholic drink (%)" = getTable1Stats("ageoffirstdrink_cat",outcome,desc_both,statistics=stats,na.rm=na.rm.var),
           "First introduced to drinking by" = getTable1Stats("introductiontodrinking",outcome,desc_both,statistics=stats,na.rm=na.rm.var),
           "Types of alcoholic drinks taken" = getTable1Stats("alcoholicdrinkstakes_cat",outcome,desc_both,statistics=stats,na.rm=na.rm.var),
           #"Types of alcoholic drinks taken" = getTable1Stats("alcoholicdrinkstakes",outcome,desc_both,statistics=stats,na.rm=na.rm.var),
           #"Other substances used" = getTable1Stats("othersubstances",outcome,desc_both,statistics=stats,na.rm=na.rm.var),
           "Other substances used" = getTable1Stats("othersubstances_cat",outcome,desc_both,statistics=stats,na.rm=na.rm.var),
          "Ever recieved care for substance related problems" =getTable1Stats("everreceivedcare",outcome,desc_median,statistics=stats,na.rm=na.rm.var)
          )%>%
   htmlTable( caption  = "<b> Alcohol and drug abuse characteristics </b>",useViewer=T,ctable = TRUE, align = 'lcccc',
              n.cgroup = c(1,2,1 ), cgroup = c('', 'Gender', '') ,tfoot="<sup>&Dagger;</sup>"
              )%>% htmlTable::addHtmlTableStyle(col.rgroup = c("#FFF", "#EEF")) 



```












```{r  cache=T, fig.width=5, warning=FALSE, eval=T}

cont_univ_df <- dataset %>% select(
              domain1_pre, domain1_post, domain2_pre, domain2_post, domain3_pre, domain3_post, domain4_pre, domain4_post,audit2,audit1,ismiscore,
              ageoffirstdrink,age
          ) %>% select_if(is.numeric) %>% dplyr::mutate(row_no =1:n() )

univ_cat_df <- dataset %>% select(-alcoholicdrinkstakes,-othersubstances)%>%
                  select_if(function(col) {is.factor(col) | is.character(col)})
```

# Alcohol Drinking Patterns

This assesses the benefits of community-based alcohol support groups on alcohol drinking patterns  of the patients in terms of number of drinking days and the number of drinks per drinking day 



## Drinking Days Per Month



```{r  align = 'center', warning=FALSE}
drinking_patterns <- readxl::read_excel("Group Evaluation-drinking patterns.xlsx")%>%filter(ID%in%c(116,12,35,83)==F)


# define ggplot alpha level
drinking_patterns <- drinking_patterns %>%
    mutate(prob = runif(n()))%>%
   mutate(alpha = ifelse(prob > quantile(prob,1/10)[[1]],0.1,.8))

# gen ling
drinking_patterns.long=drinking_patterns %>% 
  pivot_longer(
           cols = starts_with("Drinking Days"),names_to = "month",values_to = "drinking_days",values_drop_na = TRUE) %>%
            mutate(
              month=as.numeric(str_replace(month,"Drinking Days Month ","")),
              ID=as.character(ID)
            )


subset <- sample_n(drinking_patterns, 20)$ID
drinking_patterns.few = drinking_patterns.long[drinking_patterns.long$ID %in% subset,]
ggplot(drinking_patterns.few, aes(x = month, y = drinking_days)) + 
   geom_point()+
    geom_line(aes(group = ID))+
   facet_wrap(~factor(ID)) + theme_minimal()

```


The above plot represent a subset of study participants. In the grid above, each box represent trends in average drinking days per month for each partitcipant.
In general, there is a reduction in the number of drinking days per month for each participant across the 6 months period. The group level trend (regression) is depicated in the plot below:


```{r  align = 'center', warning=FALSE}

drinking_patterns.long%>%ggplot(aes(x = month, y = drinking_days)) +
  geom_jitter(aes(group = ID),color=drinking_patterns.long$ID)+
    guides(colour=FALSE) +
    labs(y = "Monthly Drinking Days", x="Month")+
     guides(fill=FALSE)+
     aes(alpha=alpha) +
    geom_smooth(se=T, colour="black", method = "lm", linetype=6)+
  	theme_minimal()+   theme(legend.position="none")+
     ggpmisc::stat_fit_glance(method = 'lm', method.args = list(formula = y ~ x),
                  geom='text', aes(label=ifelse(..p.value..< 0.001, "p<0.001**", 
                                                ifelse(..p.value..>=0.001 & ..p.value..<0.05, "p<0.05*", "p>0.05"))))




  

```



The dotted line is the cohort level regression and as expected it has a negative slope - indicating an overall reduction in drinking days per month


## Average Drinks per day Per Month



```{r  align = 'center', warning=FALSE}


# gen ling
drinking_patterns.long2=drinking_patterns %>% 
  pivot_longer(
           cols = starts_with("Drinks per day"),names_to = "month",values_to = "drinks_per_day",values_drop_na = TRUE) %>%
            mutate(
              month=as.numeric(str_replace(month,"Drinks per day Month ","")),
              ID=as.character(ID)
            )


subset <- sample_n(drinking_patterns, 20)$ID
drinking_patterns.few = drinking_patterns.long2[drinking_patterns.long2$ID %in% subset,]
ggplot(drinking_patterns.few, aes(x = month, y = drinks_per_day)) + 
   geom_point()+
    geom_line(aes(group = ID))+
   labs(y = "Average Drinks per day Per Month", x="Month")+
   facet_wrap(~factor(ID)) + theme_minimal()


```



The above plot represent a subset of study participants. In the grid above, each box represent trends in average number of drinks per day per month for each partitcipant.
For a few participants, there is a visible reduction in the number of  drinks per day per month across the 6 months period. The group level trend (regression) is depicated in the plot below:


```{r  align = 'center', warning=FALSE}

drinking_patterns.long.no = drinking_patterns.long2%>% filter(drinks_per_day < 10)
drinking_patterns.long.no%>%ggplot(aes(x = month, y = drinks_per_day)) +
  geom_jitter(aes(group = ID),color=drinking_patterns.long.no$ID)+
    guides(colour=FALSE) +
    labs(y = "Average Drinks per day Per Month", x="Month")+
     guides(fill=FALSE)+
     aes(alpha=alpha) +
    geom_smooth(se=T, colour="black", method = "lm", linetype=6)+
  	theme_minimal()+   theme(legend.position="none")+
     ggpmisc::stat_fit_glance(method = 'lm', method.args = list(formula = y ~ x),
                  geom='text', aes(label=ifelse(..p.value..< 0.001, "p<0.001**", 
                                                ifelse(..p.value..>=0.001 & ..p.value..<0.05, "p<0.05*", "p>0.05"))))
  

```



The dotted line is the group (cohort) level regression and as expected it has a negative slope - indicating an overall reduction in average number of dayly drinks per month. The reduction is not as substancial as the above (Drinking Days Per Month)

## Multilevel analysis

Please note that in our analysis we remove these participants : 116,12,35,83


```{r results="asis", align = 'center', warning=FALSE}
dataset$ID=as.character(dataset$NO)
# Data Prep
drinking_patterns.long.df= drinking_patterns.long%>%select(ID,month,drinking_days)%>% 
  inner_join(drinking_patterns.long2%>%select(ID,month,drinks_per_day), by = c("ID", "month"))%>% 
   left_join(dataset%>%select(
     ID,gender,maritalstatus ,religion,education,work,
                     income,currentfamily,supportsystem,familyalcoholhistory,introductiontodrinking,
                     everreceivedcare,age_cat,ageoffirstdrink_cat,alcoholicdrinkstakes_cat,othersubstances_cat
   ), by = c("ID"))

# = drinking_days ~ month + gender + maritalstatus +  education + income + currentfamily + familyalcoholhistory +  alcoholicdrinkstakes_cat
full.glm <- glm(drinking_days ~ month+drinks_per_day+gender+maritalstatus +religion+education+work+
                     income+currentfamily+supportsystem+familyalcoholhistory+introductiontodrinking+
                     everreceivedcare+age_cat+ageoffirstdrink_cat+alcoholicdrinkstakes_cat+othersubstances_cat
                   , data = na.omit(drinking_patterns.long.df))
#min.glm <- MASS::stepAIC(full.glm, trace = T)
min.glm= glm(drinking_days ~ month + gender + maritalstatus +  education + income + currentfamily + familyalcoholhistory +  alcoholicdrinkstakes_cat, data =drinking_patterns.long.df)
#summary(min.glm)

```

```{r results="asis", align = 'left', warning=FALSE}


m_null=lmer(drinking_days~1+(1|ID),drinking_patterns.long.df,REML=F)

m1<-lmer(drinking_days~1+(month|ID),drinking_patterns.long.df,REML=F)


m5<-lmer(drinks_per_day~month+drinking_days+gender + maritalstatus +  education + income + currentfamily + familyalcoholhistory +  alcoholicdrinkstakes_cat+
           month*drinking_days+
          (month|ID),drinking_patterns.long.df,REML=F)

m6<-lmer(drinking_days~month+drinks_per_day+gender + maritalstatus +  education + income + currentfamily + familyalcoholhistory +  alcoholicdrinkstakes_cat+
           month*drinks_per_day+
          (month|ID),drinking_patterns.long.df,REML=F)


m7<-lmer(drinking_days~month+drinks_per_day+gender + maritalstatus +
           month*drinks_per_day+  month*gender +  month*maritalstatus +
          (month|ID),drinking_patterns.long.df,REML=F)
#summary(m5)
summ(m6, confint = TRUE, pvals=T)
jtools::plot_summs(m6,  coefs = c(
 'month' =  'month',
'drinks per day' ='drinks_per_day',
'genderfemale' ='genderfemale',
'marital status - single'	 ='maritalstatussingle'	,
'marital status - separated' = 'maritalstatusseparated' ,
'education - elementary/primary'	 = 'educationelementary/primary',
'education - secondary' ='educationsecondary', 
'education - college/university' ='educationcollege/university',
'income - 2000-10000' ='income2000-10000',
'income - 11000-20000' ='income11000-20000',
'income - 21000-50000' ='income21000-50000',
'current family - spouse'	 ='currentfamilyspouse',
'current family - parents' = 'currentfamilyparents' ,
'current family - siblings'	 = 'currentfamilysiblings',
'current family - other' = 'currentfamilyother',
'family alcohol history - no' = 'familyalcoholhistoryno',
'alcoholic drinks takes busaa'  = 'alcoholicdrinkstakes_catbusaa',
'alcoholic drinks takes - chang’aa' = 'alcoholicdrinkstakes_catchang’aa',
'alcoholic drinks takes - none' = 'alcoholicdrinkstakes_catnone',
'alcoholic drinks takes - others' = 'alcoholicdrinkstakes_catothers',
'month:drinks per day' = 'month:drinks_per_day'
))


texreg::htmlreg(list(m5, m6),

          single.row = T,
          #stars = numeric(0),
          title = "dddddd",
          custom.note = "P-values Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1")

stargazer(as(m7,"lmerMod"), as(m7,"lmerMod"), type = "html",
          title="An Illustrative Model Using Cake Data", dep.var.labels.include = FALSE
          #covariate.labels=c()
)

```


# QOL Bivariate Analysis




 
##  QOL before and after support group (SG)

This establishes the benefits of community-based alcohol support groups on the perceived quality of life of patients with alcohol and drug abuse disorders.

```{r  align = 'center', warning=FALSE}

domain1<- dataset %>% pivot_longer(
           cols = starts_with("domain1_p"),names_to = "domain_label",values_to = "domain_value",values_drop_na = TRUE)%>%
                  mutate(
                    domain_name = factor(ifelse(domain_label=="domain1_post", "Post SG","Prior SG"), 
                                         levels = c( "Prior SG","Post SG"))
                  )
domain2<- dataset %>% pivot_longer(
           cols = starts_with("domain2_p"),names_to = "domain_label",values_to = "domain_value",values_drop_na = TRUE)%>%
                  mutate(
                    domain_name = factor(ifelse(domain_label=="domain2_post", "Post SG","Prior SG"), 
                                         levels = c( "Prior SG","Post SG"))
                  )
domain3<- dataset %>% pivot_longer(
           cols = starts_with("domain3_p"),names_to = "domain_label",values_to = "domain_value",values_drop_na = TRUE)%>%
                  mutate(
                    domain_name = factor(ifelse(domain_label=="domain3_post", "Post SG","Prior SG"), 
                                         levels = c( "Prior SG","Post SG"))
                  )
domain4<- dataset %>% pivot_longer(
           cols = starts_with("domain4_p"),names_to = "domain_label",values_to = "domain_value",values_drop_na = TRUE)%>%
                  mutate(
                    domain_name = factor(ifelse(domain_label=="domain4_post", "Post SG","Prior SG"), 
                                         levels = c( "Prior SG","Post SG"))
                  )
na.rm.var=F;stats=T;outcome="domain_name";
# Getting descriptive statistics 
mergeDesc("Domain1 | Physical health" = getTable1Stats("domain_value","domain_name",desc_both,data=domain1,statistics=stats,na.rm=na.rm.var, header_count = F ),
          "Domain2 | Psychological" = getTable1Stats("domain_value","domain_name",desc_both,data=domain2,statistics=stats,na.rm=na.rm.var, header_count = F ),
          "Domain3 | Social Relationship" = getTable1Stats("domain_value","domain_name",desc_both,data=domain3,statistics=stats,na.rm=na.rm.var, header_count = F ),
          "Domain4 | Environment" = getTable1Stats("domain_value","domain_name",desc_both,data=domain4,statistics=stats,na.rm=na.rm.var, header_count = F )
         
          ) %>%
   htmlTable( caption  = "<b> QOL before and after support group (SG) </b>",useViewer=T,ctable = TRUE, align = 'lcccc',
              n.cgroup = c(1,2,1 ), cgroup = c('', 'Quality of Life', '') ,tfoot="<sup>&Dagger; TODO: Please delete total column </sup>"
              )%>% htmlTable::addHtmlTableStyle(col.rgroup = c("#FFF", "#EEF")) 
```




We also plotted box-plot for each of the variables to do a visual comparison of QOL between pre and post intervention.



```{r  cache=T, warning=FALSE}
angle=10;size = 0.8; jitter = 0.3;
  value="domain_name";
   p1<- dataset %>% pivot_longer(
           cols = starts_with("domain1_p"),names_to = "domain_label",values_to = "domain_value",values_drop_na = TRUE )%>%
                  mutate(
                    domain_name = factor(ifelse(domain_label=="domain1_post", "Post Intervention","Prior Intervention"), 
                                         levels = c( "Prior Intervention","Post Intervention"))
                  )%>%
           ggboxplot(x ="domain_name", y = "domain_value",merge = TRUE,rug = TRUE ,   color = value, palette = "jco", add = "jitter", shape = value,
                      add.params = list(size = size, jitter = jitter),  repel = TRUE, subtitle = "Domain1 | Physical health ")+
                stat_compare_means()+theme_minimal() +theme(axis.title = element_blank())

   
   p2<- dataset %>% pivot_longer(
           cols = starts_with("domain2_p"),names_to = "domain_label",values_to = "domain_value",values_drop_na = TRUE )%>%
                  mutate(
                    domain_name = factor(ifelse(domain_label=="domain2_post", "Post Intervention","Prior Intervention"), 
                                         levels = c( "Prior Intervention","Post Intervention"))
                  )%>%
           ggboxplot(x ="domain_name", y = "domain_value",merge = TRUE, color = value, palette = "jco", add = "jitter", shape = value,
                      add.params = list(size = size, jitter = jitter),  repel = TRUE, subtitle = "Domain2 |  Psychological")+
                stat_compare_means()+ theme_minimal() +theme(axis.title = element_blank())
   
   p3<- dataset %>% pivot_longer(
           cols = starts_with("domain3_p"),names_to = "domain_label",values_to = "domain_value",values_drop_na = TRUE )%>%
                  mutate(
                    domain_name = factor(ifelse(domain_label=="domain3_post", "Post Intervention","Prior Intervention"), 
                                         levels = c( "Prior Intervention","Post Intervention"))
                  )%>%
           ggboxplot(x ="domain_name", y = "domain_value",merge = TRUE, color = value, palette = "jco", add = "jitter", shape = value,
                      add.params = list(size = size, jitter = jitter),  repel = TRUE, subtitle = "Domain3 |  Social Relationship")+
                stat_compare_means()+ theme_minimal() +theme(axis.title = element_blank())
   
   
   p4<-  dataset %>% pivot_longer(
           cols = starts_with("domain4_p"),names_to = "domain_label",values_to = "domain_value",values_drop_na = TRUE )%>%
                  mutate(
                    domain_name = factor(ifelse(domain_label=="domain4_post", "Post Intervention","Prior Intervention"), 
                                         levels = c( "Prior Intervention","Post Intervention"))
                  )%>%
           ggboxplot(x ="domain_name", y = "domain_value",merge = TRUE, color = value, palette = "jco", add = "jitter", shape = value,
                      add.params = list(size = size, jitter = jitter),  repel = TRUE, subtitle = "Domain4 | Environment")+
                stat_compare_means()+ theme_minimal() +theme(axis.title = element_blank())
   
  legend = gtable_filter(ggplot_gtable(ggplot_build(p1 + theme(legend.position="right")+
  theme(legend.title=element_blank()))), "guide-box")

  
 #grid.arrange(legend)
 grid.arrange( 
   
   p1 + theme(legend.position="none"),
   p2 + theme(legend.position="none"),
   p3 + theme(legend.position="none"),
   p4 + theme(legend.position="none")
               ) 
 


```


clearly we are able to see a significant and substancial difference between prio and post support group intervention across the 4 domains


##  Difference in QOL before and after support group (SG)

To assess wheather there is any significant association between QOL difference (post - pre)  and  socio demographics  / baseline drinking characteristics across the 4 domains


```{r  results="asis", cache=T, warning=FALSE}

dataset$domain1_diff= dataset$domain1_post - dataset$domain1_pre 
dataset$domain2_diff= dataset$domain2_post - dataset$domain2_pre
dataset$domain3_diff= dataset$domain3_post - dataset$domain3_pre
dataset$domain4_diff= dataset$domain4_post - dataset$domain4_pre

angle=10;na.rm.var=T;stats=T;outcome="gender";
for (value in  colnames(univ_cat_df))  { 
  #print(value)
  
   n = length(unique(dataset[[value]]))
  if(value=="maritalstatus"||value=="religion"||value=="income"){
    n=n+1
  }
 
  if(value=="ageoffirstdrink_cat"){
    n=n-1
  }
 print(mergeDesc(
            
             "Domain1 | Physical health (post - pre)" = getTable1Stats("domain1_diff",value,desc_mean,statistics=stats,na.rm=na.rm.var),
             "Domain2 | Psychological (post - pre)" = getTable1Stats("domain2_diff",value,desc_mean,statistics=stats,na.rm=na.rm.var),
             "Domain3 | Social Relationship (post - pre)" = getTable1Stats("domain3_diff",value,desc_mean,statistics=stats,na.rm=na.rm.var),
             "Domain4 | Environment (post - pre)" = getTable1Stats("domain4_diff",value,desc_mean,statistics=stats,na.rm=na.rm.var)
            
           
            )%>%
     htmlTable( caption  = paste0("Domain Comparison | ",value),useViewer=T,ctable = TRUE, align = 'lcccc',
                n.cgroup = c(1,n,1 ), cgroup = c('', value, '') ,tfoot="<sup>&Dagger;</sup>"
                )%>% htmlTable::addHtmlTableStyle(col.rgroup = c("#FFF", "#EEF")) )

}

```









